### 2024-07-09 manuscript/appendix.Rmd
appendix up to date

---
title: |
     Prey movement shapes the development of predator expertise in a virtual bi-trophic system: \
     Appendix

output: 
  word_document:
    reference_docx: appendix-styles.docx
    fig_caption: yes
csl: chicago-author-date.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(data.table)
library(brms)
library(flextable)
library(officer)
library(dplyr)
```

\newpage

```{r tableprep, include = FALSE}
# ===========================================================================
# 1. Extract the fixed effects from the MV model
# ===========================================================================

# Prepare the table ---------------------------------------------------------

# Model fit
fit <- readRDS(
  here::here("outputs", "outputs_models", "DHMLM-RandomSample.rds")
)

# Fixed effects table
results <- data.table(
  round(summary(fit, robust = TRUE, prob = 0.89)$fixed, digits = 2)[, 1:4],
  keep.rownames = TRUE
)
setnames(results, "rn", "parameter")

# Trait variable (to remove after)
results[parameter %like% "predspeed", trait := "predator speed"][
  parameter %like% "preyavgspeed", trait := "prey speed"][
    parameter %like% "success", trait := "hunting success"
  ]

# mu parameters table
tab <- results[
    !(parameter %like% "sigma"),
    c(6, 1, 2, 4, 5)
]
tab <- tab[trait == "hunting success"]

# Rename parameter values ---------------------------------------------------

# Rename
tab[,
  parameter := 
      c("game duration",
        "cumulative experience",
        "group 1",
        "group 4",
        "group 2",
        "group 3")
]
# Reorder rows
tab <- tab[
  c(1,2,3,5,6,4)
]
tab[, trait := NULL]

# ===========================================================================
# ===========================================================================

# ===========================================================================
# 2. Compute the table using flextable
# ===========================================================================

# Prepare the table parameters ----------------------------------------------

# Custom header
my_header <- data.frame(
   col_keys = c("parameter",
                "Estimate",
                "l-89% CI",
                "u-89% CI"),
   line1 = c("Parameter",
             "Estimate",
             "lower 89% CI",
             "upper 89% CI"),
   stringsAsFactors = FALSE
)

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "autofit")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
   x <- width(x, j = 1, width = 3.78, unit = "cm")
   x <- width(x, j = c(3:4), width = 3.3, unit = "cm")
}
```

```{r tableS1, ft.align = "left"}
# Create the table ----------------------------------------------------------

tabf <- tab %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  align(align = "left", part = "all", j = 1) %>%
  align(align = "center", part = "all", j = c(2:4)) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e. group in the main text)"
    ),
    ref_symbols = " "
  ) %>%
  set_caption(
    tab,
    caption = "Table S1. Fixed effects table comparing the success of random groups of players with different amounts of matches played to the group presented in the main text.",
    style = "Table Caption"
  ) %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer") %>%
  font(fontname = "Times New Roman", part = "all")

tabf

# ===========================================================================
# ===========================================================================
```

\newpage

```{r figS1, out.width = "100%", fig.cap = "Figure S1. Posterior distributions of the median hunting success per group of players. The values were back transformed to probability scale and can be interpreted as the probability of capturing the four prey. Group 1: <50 matches, Group 2: between 50 and 99 matches, Group 3: between 100 and 299 matches, Group 4: > 299 (i.e. group in the main text)."}

knitr::include_graphics(
  here::here(
    "outputs", "outputs_figures",
    "figureS1.png"
  )
)
```